---
title: 'Local Demographic Changes and US Presidential Voting, 2012-2016: A Replication'
author: "Debi Gonzalez"
date: "4/25/2020"
output: bookdown::pdf_document2
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


library(haven)
library(tinytex)
library(stargazer)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(gt)
library(broom)
library(cowplot)
library(plm)

```


```{r data setup}

# Here I load two of the original datsets, which I will use to replicate the main
# results of Hopkins, Hilla, and Huber.

geo_scope <- read_dta("dataverse_files/HHHDemogThreatRepArchive/GeographicScopeData.dta")
  
precinct_data <- read_dta("dataverse_files/HHHDemogThreatRepArchive/PrecinctData.dta")

```

\newpage
# Abstract

Hilla, Hopkins, and Huberc (2019) argue that demographic changes are not associated with increased Republican vote share at the precinct level between 2012 and 2016. I succesfully replicated Hilla, Hopkins and Huberc’s results, except for minor discrepancies which do not affect their conclusions. A robustness test using a subset analysis focusing on the state of FL indicates, despite minor state-specific deviance and poor statistical significance due to reduced sample size, that the authors' overall findings are valid. In addition, I constructed a fixed effects model by state, which further confirmed original findings even under fixed effects assumptions. In a twist of analysis, I constructed a new model analyzing the association between Republican vote share and economic indicators while controlling for Hispanic proportion. The latter models appear to indicate a stronger relationship between unemployment proportion change and GOP vote share than between changes in Hispanic proportion and GOP vote share, which suggests an opportunity for further research. 



# Introduction

Immigration and demographic changes have become salient in American political discourse in great part due to Trump's capaign for tthe 2016 election. Inspired by previous inconclusive research on threatened response and county-level electoral outcomes, Hilla, Hopkins, and Huberc, henceforth HHH, compiled a novel, precinct-level data set of election results and census demographic measures for almost 32,000 precincts across 7 states to explore how demographic changes influenced republican vote share in affected precincts across the United States. The authors employed regression models with varying explanatory variables and measures of demographic change. Through this analysis, the authors find consistent evidence that 'influxes of Hispanics or non-citizen immigrants' were associated with decreased Republican vote share, which was assumed by the authors to mean shifts towards Trump's opponent in 2016. Dan Hopkins offers an overview and a link to the data for this research paper on [his website](http://web.sas.upenn.edu/danhop/research/">his website). The full original paper can be found [here](https://www-pnas-org.ezp-prod1.hul.harvard.edu/content/pnas/116/50/25023.full.pdf).


Using R and RStudio, I succesfully replicated Hilla, Hopkins and Huberc’s results, except for minor discrepancies which do not affect their conclusions [@R]. Some of these discepancies can be associated with the inherent assumptions of the Stata software versus R. The replication process was lengthy and focused mainly on translating local variables from Stata (the coding language in which the original analysis was conducted) into R using the original datasets to construct the main eight models supporting the argument. A step by step replication of the 8 main models and graphics of the paper are included in the appendix. Thanks to the generosity of the authors the code and data used in this project are available at [Harvard Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/J5GCZQ). For more information about this project please visit my project's Github repo^[[The data and code for this project can be found here.](https://github.com/deboragonzalez/gop_voters_hisp_changes)].
 
The first paragraph is a review of the paper you are replicating. Place it within the relevant literature, via a key citation or two. Highlight implications and caveats. 

The second paragraph provides more details on your replications. Mention that you used R, and provide an citation to R which will appear in your bibliography. (See citation().) Cite the location from which you got the data and code for your replication. (This might be to the Dataverse, a webpage or “personal communication” with the author.) Provide a footnote with a link to your repo.

The Third and Fourth paragraphs: What did you do? What did you find?

\newpage
# Paper & Literature Review


Using a data set of nearly 32000 precincts across the states of Florida, Georgia, Michigan, Nevada, Ohio, Pennsylvania, and Washington, HHH aimed to understand whether demographic changes at low levels of aggregation were associated with vote shifts toward an anti-immigration presidential candidate between 2012 and 2016.

if something is worth understanding about the literature
any relevant literature that has come out since then
provide a summary in whatever level of detail you think is appropriate. 


# Replication


# Robustness Check #1: Subsetting the Data


# Robutness Check #2: Fixed Effects Model


# A New Model: Economic Indicators & Hispanic Population Change


# Conclusion

The first paragraph is a review of the paper you are replicating. Place it within the relevant literature, via a key citation or two. Highlight implications and caveats. 

The second paragraph provides more details on your replications.

Conclusion, the last paragraph is where you get to speculate: What does it all mean? What should we research next?

# Appendix

# References

